Enabling conditions for effective community based marine reserves in small scale fisheries

### 2018-02-21 Frontiers_Manuscript/Villasenor-Derbez.Rmd
Added abstract to the paper

---
title: "Enabling conditions for effective community based marine reserves in small scale fisheries"
rtitle: "Marine reserve effectiveness"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
institution: "
$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico
"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "jvillasenor@bren.ucsb.edu"
abstract: "Abstract after comments from co-authors"
keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico, Social--Ecological Systems"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

>>>>

---
title: "Enabling conditions for effective community based marine reserves in small scale fisheries"
rtitle: "Marine reserve effectiveness"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
institution: "
$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico
"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "jvillasenor@bren.ucsb.edu"
abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. However, overfishing and unsustainable fishing practices threaten the marine environment and jeopardize the wellbeing of coastal communities. A common approach to protect the environment and recover overexploited stocks is to implement no-take marine reserves (areas where all extractive activities are off-limits). In small-scale fisheries, these are sometimes implemented as community-based reserves, where a group of fishers collectively agree to close an area to fishing. While we know that reductions in fishing effort are followed by a series of ecological benefits (increased biomass, abundance, and species diversity), we do not fully understand how environmental and governance dynamics influence the conservation and fisheries benefits of community-based marine reserves. In this work, we evaluate the ecological outcomes of four reserves established by three coastal communities in temperate and tropical ecosystems of Mexico. By combining causal inference techniques with an operationalization of the social-ecological systems framework, we identify the environmental and social conditions that enable reserve effectiveness. Our results show a strong interaction between environmental variation and community organization, which influences reserve effectiveness. For example, the most effective reserve had strong governance structures accompanied with low environmental variability. Thus, even when a community is well organized (and reserves are well enforced), environmental variation can hinder the benefits of a reserve, and vice versa. Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making."
keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico, Social--Ecological Systems"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

### 2018-02-21 Frontiers_Manuscript/Villasenor-Derbez.tex
Added abstract to the paper

Abstract after comments from co-authors

>>>>

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. However, overfishing and
unsustainable fishing practices threaten the marine environment and
jeopardize the wellbeing of coastal communities. A common approach to
protect the environment and recover overexploited stocks is to implement
no-take marine reserves (areas where all extractive activities are
off-limits). In small-scale fisheries, these are sometimes implemented
as community-based reserves, where a group of fishers collectively agree
to close an area to fishing. While we know that reductions in fishing
effort are followed by a series of ecological benefits (increased
biomass, abundance, and species diversity), we do not fully understand
how environmental and governance dynamics influence the conservation and
fisheries benefits of community-based marine reserves. In this work, we
evaluate the ecological outcomes of four reserves established by three
coastal communities in temperate and tropical ecosystems of Mexico. By
combining causal inference techniques with an operationalization of the
social-ecological systems framework, we identify the environmental and
social conditions that enable reserve effectiveness. Our results show a
strong interaction between environmental variation and community
organization, which influences reserve effectiveness. For example, the
most effective reserve had strong governance structures accompanied with
low environmental variability. Thus, even when a community is well
organized (and reserves are well enforced), environmental variation can
hinder the benefits of a reserve, and vice versa. Our results are
particularly relevant under present changing climate conditions, as they
can better inform management and decision making.
